---
title: "Presentation_2dot0"
author: "Fabian Oberreiter, Konrad Medicus, Tobias Hilgart"
date: "13 5 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Motivation

Let's take a look at the following bivariate sample:

```{r}
n = 50
x = rnorm(n,1,1)
y = sin(x) 
plot(x,y)
```
which we obtained from $X \sim \mathcal(N)(1,1)$ and $Y = \sin(X)$.

Would it be easier to predict $Y$ given $X$, or $X$ given $Y$?


## Motivation cont.

So, we saw how we could have an asymmetrical dependence structure. There are actual examples for that in nature as well:

<!-- Vielleicht finden wir ein anderes Beispiel  --> 

## Motivation cont.

So, why could this be a problem?

How would you usually try to quantify dependence?

Correlation comes to mind.

Let's try it:

```{r, echo=TRUE}
cor(x,y)
cor(y,x)
```
 

## Motivation cont.
We got the same value. Indeed, by the Definition of correlation (or covariance), we will always get $r(X,Y) = r(Y,X)$. The same happens, if we calculated the spearman or kendall correlation instead (same problem).

It seems, that correlation is not applicable to measure asymmetric dependence.

We would like to have a dependency measure, that can also classify asymmetric dependency. How could we go about obtaining/constructing one?

## Motivation cont.

What would we like for such a dependency measure, say *q*?

* $q(X,Y) \in [0,1]$

* $q(X,Y) = 0$ if and only if $X$ and $Y$ are independent (independence)

* $q(X,Y) = 1$ if and only if $Y$ is a function of $X$ (complete dependence)

We want a higher *q(X,Y)*, if there is a higher dependency of $Y$ on $X$. It may well be, that $q(X,Y) \neq q(Y,X)$.

Additionally, scale changes should not affect the outcome.

## Outline of the qad-measure's construction:

# (Empirical) Copula
<!-- Wirklich sehr kurz und um die Aufgabe 1 uebernehmen zu können  -->
## emp_c_copula
## plot_density

## Exercise 1

# The dependency measure $\zeta_1$
<!-- Das sollten wir in etwa so uebernehmen koennen  --> 

## qad
## plot.qad
## predict.qad

## Exercise 2
<!-- Unsere Aufabe 2 wird aufgrund der Motivation jetzt hinfaellig, wir ziehen am besten 3 vor  --> 

## pairwise
## heatmap

## Exercise 3

# Testing
<!-- Sollten wir vermutlich viel ausfuehrlicher machen, vlt eine Aufgabe, den p-Wert zu Fuß auszurechnen?  --> 

## cci

## Exercise 4
<!-- Vielleicht an RTR Aufgabe anknuepfen, auf Unabhaengikeit testen?  --> 

